---
title: "House prices during Covid"
output: html_notebook
---

```{r,message=FALSE}
# load packages
library(tidyverse)
library(astsa)
library(forecast)
library(lubridate)
```

# Part 1 Time series analysis (alternative)
# load data
```{r}
us_hv<- read_csv("country_level_hv.csv")

us_hv%>%filter(year_month >= '2018-01-01') -> us_hv

us_sfr_ts <- ts(us_hv$value_sfr, start =c(2018, 1), frequency = 12)

plot.ts(us_sfr_ts)
```

# acf & pacf plots of the series
```{r}
ggAcf(us_sfr_ts)
ggPacf(us_sfr_ts)
```


```{r}
us_sfr_ts_auto <- auto.arima(us_sfr_ts)
us_sfr_ts_auto
checkresiduals(us_sfr_ts_auto)

```
```{r}
us_sfr_ts_fit1 <-arima(us_sfr_ts, order=c(1, 2, 1))
summary(us_sfr_ts_fit1)

us_sfr_ts_fit1%>%forecast(h = 10)%>%autoplot()

(302307.5 - 265679)/265679*100
```

```{r}
us_sfr_ts_fit2<- holt(us_sfr_ts)
summary(us_sfr_ts_fit2)
holt(us_sfr_ts)%>%autoplot()
```

# other predictors
```{r}
us_hv%>%mutate(is_covid = if_else(year_month > '2020-04-01', 1, 0)) -> us_hv

us_sfr_ts_fit3 <- arima(us_sfr_ts, order= c(1,2,1), xreg = us_hv$is_covid)
summary(us_sfr_ts_fit3)
checkresiduals(us_sfr_ts_fit3)
	
```

# Part 2 regression
# Considering the fact that the arima model didn't perform well in this case, we tried to construct regression models to explore some factors influencing the house price during covid

# load data
```{r}
# load model related data 
# those are some factors we found during the exploratory analysis that would potentially affecting the house price 

# factors including state level house prices, state level covid cases and deaths, state level total number of vaccinations, nationwide mortage rate, and nationwide total number of house completion

state_hv<- read_csv("state_level_hv.csv")
state_case<- read_csv("United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
state_vaccination<- read_csv("us_state_vaccinations.csv")
mortgage_rate <- read_csv('mortgage_rate.csv') %>% clean_names()
house_completion <- read_csv('house_completion.csv') %>% clean_names()
population <- read_csv("population.csv")%>%select(NAME, POPESTIMATE2019)

```

```{r}
cor(state_case$new_case, state_case$new_death)

```

# data clearning
```{r}

state_case%>%
  mutate(submission_date = mdy(submission_date),
         year = year(submission_date),
         month = month(submission_date))%>%
  group_by(state, year, month)%>%
  summarise(tot_cases_month = max(new_case),
            tot_deaths_month = max(new_death)) ->state_case_clean

# take the difference to calculate the monthly growth in total number of people fully vaccinated
state_vaccination%>%
  mutate(year = year(date),
         month = month(date))%>%
  group_by(location, year, month)%>%
  summarise(people_fully_vaccinated_per_hundred_diff = max(people_fully_vaccinated_per_hundred, na.rm = TRUE)-min(people_fully_vaccinated_per_hundred, na.rm = TRUE))%>%
  select(location, year, month, people_fully_vaccinated_per_hundred_diff)-> state_vaccination_clean

# calculate the monthly growth rate of single-family homes
# from 2019 Feburary
state_hv%>%
  group_by(StateName)%>%
  mutate(num_listing_sfr_lag = lag(num_listing_sfr),
         num_listing_sfr_growth_rate = (num_listing_sfr - num_listing_sfr_lag)/num_listing_sfr_lag*100)%>%
  filter(year_month > '2019-02-01')%>%
  mutate(year = year(year_month),
         month = month(year_month))%>%
  select(RegionName, StateName, year_month, year, month, value_sfr_growth, num_listing_sfr_growth_rate)-> state_hv_clean

# nationwide mortgage rate
# from 2019 Feburary
mortgage_rate %>%
  mutate(time=mdy(time)) %>%
  filter(time > '2019-02-01') %>%
  rename(date=time, mortgage_rate=usa) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  subset(select=-c(ym, date)) -> mortgage_rate_clean

# nationwide total number of house completion
# from 2019 Feburary
house_completion %>%
  mutate(date = ymd(paste(ym,'-02',sep=''))) %>%
  filter(date > '2019-02-01') %>%
  rename(house_completion=usa) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  subset(select=-c(ym, date)) -> house_completion_clean


# join cleaned datasets
state_hv_clean%>%
  left_join(state_case_clean, by = c("StateName" = "state", "year", "month"))%>%
  left_join(state_vaccination_clean, by = c("RegionName"="location", "year", "month")) %>%
  left_join(mortgage_rate_clean, by = c("year" , "month")) %>%
  left_join(house_completion_clean, by = c("year" , "month")) -> train_df

# replace na with 0
train_df[is.na(train_df)] <- 0

# considering that the total number of covid cases and deaths are highly affected by the population, calculate the total cases and deaths per thousand
train_df%>%
  left_join(population, by = c("RegionName" = "NAME"))%>%
  mutate(tot_cases_per_thousand = tot_cases_month/POPESTIMATE2019*1000,
         tot_deaths_per_thousand = tot_deaths_month/POPESTIMATE2019*1000)%>%
  mutate(across(.cols = is.numeric, round, 2)) -> train_df


```



# regression model 1

```{r}
lm_fit1 <- lm(value_sfr_growth ~ num_listing_sfr_growth_rate + 
                new_death_growth_rate  + 
                people_fully_vaccinated_per_hundred_diff + 
                mortgage_rate + 
                house_completion, data = train_df)

summary(lm_fit1)

```
# regression model 2

```{r}
# add a binary variable indication if the datapoint is during covid
train_df<- train_df %>% mutate(in_covid = if_else(year_month > "2020-03-01", 1, 0))

# get rid of some insignificant variables
lm_fit2 <- lm(value_sfr_growth ~ num_listing_sfr_growth_rate + 
                in_covid  + 
                people_fully_vaccinated_per_hundred_diff, data = train_df)
summary(lm_fit2)

# R squared 0.3825
```

# regression model 3

```{r}

lm_fit3 <- lm(value_sfr_growth ~ num_listing_sfr_growth_rate + 
                tot_deaths_per_thousand + 
                people_fully_vaccinated_per_hundred_diff, data = train_df)
summary(lm_fit3)

# R squared 0.218
```

# regression model 4

```{r}
# add two country level variables
lm_fit4 <- lm(value_sfr_growth ~ num_listing_sfr_growth_rate + 
                in_covid  + 
                people_fully_vaccinated_per_hundred_diff + 
                mortgage_rate + 
                house_completion, data = train_df2)

summary(lm_fit4)

# R squared 0.4341
```

